---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
date: "2025-05-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# clear env
```{r}
rm(list = ls())
```

## Libraries
```{r}
library(tidyverse)
library(here)

```

# create a new folder to save the data
```{r}
# # Create a new folder named "Data" in your working directory
# dir.create(here("Data"), showWarnings = FALSE)
# 
# # Define the url and destination file path
# url = "https://data.lacity.org/resource/9w5z-rg2h.csv"
# dest_file = here("Data/lacity_data.csv")
# 
# # Download the CSV file to the new folder
# download.file(url, destfile = dest_file, mode = "wb")
# 
# # Read the CSV file from the saved location
# data = read.csv(dest_file)
# 
# # Preview the data
# head(data)

```
### The HTTP address did not produe all the entries , so downloaded the data manually
# Can not download the data from the 2nd link provided in the email

# read data
```{r}
bldg_safty_insp_data = read_csv(here("Data", "Building_and_Safety_Inspections_20250505.csv")) %>% 
  # clean names
  janitor::clean_names()
``` 


```{r}
bldg_safty_insp_data %>% head(50)
```


## Count the unique number of entries in each column
```{r}
bldg_safty_insp_data %>% 
  summarise(across(everything(), ~ n_distinct(.)))
```
# convert to factors if entries are below a threshold
```{r}
bldg_safty_insp_data %>% 
  mutate(across(where(~ is.character(.) && n_distinct(.) < 100), as.factor)) %>% 
  head()
```

# Question 1

## Summary Table for Permits vs Inspection results
```{r}
summary_table = bldg_safty_insp_data %>%
  count(`permit_status`, `inspection_result`, sort = TRUE, name = "count")

# Display the table
summary_table
```

```{r}
bldg_safty_insp_data %>%
  count(`permit_status`, `inspection_result`, sort = TRUE, name = "count") %>% 
  mutate(count_log = log1p(count)) %>% 
# ggplot heatmap
ggplot( aes(x = `inspection_result`, y = `permit_status`, fill = count_log)) +
  geom_tile(color = "white") +
  # geom_text(aes(label = count), size = 3) +
  scale_fill_viridis_c(option = "C", name = "Log(count + 1)") +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  ) +
  labs(
    title = "Heatmap of Permit Status vs Inspection Result",
    x = "Inspection Result",
    y = "Permit Status"
  )
```

## for easy visualization and meaningful observations remove counts less than 100
```{r, fig.width=7, fig.height=8}

bldg_safty_insp_data %>%
  count(`permit_status`, `inspection_result`, sort = TRUE, name = "count") %>% 
  filter(count >100) %>% 
  mutate(count_log = log1p(count)) %>% 
# ggplot heatmap
ggplot( aes(y = `inspection_result`, x = `permit_status`, fill = count_log)) +
  geom_tile(color = "white") +
  # geom_text(aes(label = count), size = 3) +
  scale_fill_viridis_c(option = "C", direction = -1,  name = "Log(count + 1)") +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  ) +
  labs(
    title = "Heatmap of Permit Status vs Inspection Result",
    y = "Inspection Result",
    x = "Permit Status"
  )
```

# A new link for 2nd dataset is provided

## downloaded this data manually

### Read permit data in R
```{r}
bldg_permit_data = read_csv(here("Data", "Building_Permits_20250505.csv")) %>% 
  # clean names
  janitor::clean_names()
```

## 
```{r}
bldg_permit_data %>%  head()
```


## To check if the two datasets have some common addresses based on address columns

```{r}
# Find common rows based on 'permit' and 'address'
bldg_safty_insp_permit = inner_join(bldg_safty_insp_data, bldg_permit_data,
                          by = c("address"= "applicant_address_1")
                          )

print(bldg_safty_insp_permit)
```
## Some address have more than 2 entries and thats why we are getting more than 1 row for the same address there are more rows than the permit dataset
We can now select the columns we are interested in and save the data
```{r}
bldg_safty_insp_permit_2 = bldg_safty_insp_permit %>% 
  select(
    address,
    permit_status,
    inspection_type,
    inspection_result,
    status,
    permit_type,
    contractor_city,
    contractor_state,
    applicant_address_3,
    zone 
    ) %>% 
  # remove duplicates
  distinct()
bldg_safty_insp_permit_2 %>%  dim()
```

```{r}
bldg_safty_insp_permit_2 %>% head(15)
```

## I notices that LOS ANGELES had multiple entries eg LOS ANGELES, CA, LOS ANGELES ,CA, Los Angeles, L.A., CA, etc.
## The code below is to unify it
```{r}
bldg_safty_insp_permit_2 = bldg_safty_insp_permit_2 %>%
  mutate(
    applicant_address_3 = str_to_upper(applicant_address_3),                  # Make all uppercase
    applicant_address_3 = str_replace_all(applicant_address_3, "[[:punct:]]", ""), # Remove punctuation
    applicant_address_3 = str_squish(applicant_address_3),                   # Remove extra spaces
    applicant_address_3 = case_when(
      str_detect(applicant_address_3, "LOS ANGELES") ~ "LOS ANGELES, CA",
      str_detect(applicant_address_3, "L A") ~ "LOS ANGELES, CA",
      str_detect(applicant_address_3, "L\\.A") ~ "LOS ANGELES, CA",
      str_detect(applicant_address_3, "LA CA") ~ "LOS ANGELES, CA",
      str_detect(applicant_address_3, "LACA ") ~ "LOS ANGELES, CA",
      is.na(applicant_address_3) ~ "UNKNOWN",
      TRUE ~ applicant_address_3
    )
  )
```


## To get an idea of number of inspection by geography we can use applicant_address_3 which is the city
## and we can also use zone

```{r}
# Count inspections by applicant_address_3
inspection_freq = bldg_safty_insp_permit_2 %>%
  count(applicant_address_3, name = "inspection_count") %>%
  arrange(desc(inspection_count))

print(inspection_freq)

```

## PLot for inspection frequency for top 20
```{r}
inspection_freq %>% 
  # filter top 20 rows
  slice_head(n = 20) %>% 
  # Make names as factors to avoid alphabetical ordering
  mutate(applicant_address_3 = factor(applicant_address_3, levels = inspection_freq$applicant_address_3)) %>%
  ggplot(aes(x = applicant_address_3,  y = inspection_count)) +
  geom_col(fill = "blue", alpha = 0.5) +
  labs(title = "Number of Inspections by Applicant Address",
       x = "Applicant Address", y = "Inspection Count") +
  # rotate x-axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```
## Lets try the same thing by the zone
```{r}
# Count inspections by zone
inspection_freq_zone = bldg_safty_insp_permit_2 %>%
  count(zone, name = "inspection_count") %>%
  arrange(desc(inspection_count))
print(inspection_freq_zone)
```
## PLot for inspection frequency for top 20
```{r}
inspection_freq_zone_plot = 
  inspection_freq_zone %>% 
  # filter top 20 rows
  slice_head(n = 20) %>% 
  # Make names as factors to avoid alphabetical ordering
  mutate(zone = factor(zone, levels = inspection_freq_zone$zone)) %>%
  ggplot(aes(x = zone,  y = inspection_count)) +
  geom_col(fill = "blue", alpha = 0.5) +
  labs(title = "Number of Inspections by Zone",
       x = "Zone", y = "Inspection Count") +
  # rotate x-axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
print(inspection_freq_zone_plot)
```

## Inspection result by applicant address, considered as geography
```{r}
inspection_result_tbl = bldg_safty_insp_permit_2 %>%
  count(applicant_address_3, inspection_result, name = "count") %>% 
  # order by inspection result
  arrange(desc(count))

# View the table
print(inspection_result_tbl)
```



## for easy visualization and meaningful observations remove counts less than 200
```{r, fig.width=12, fig.height=7}
inspection_result_tbl %>%
  filter(count > 200) %>%
  mutate(
    count_log = log1p(count),
    applicant_address_3 = ifelse(is.na(applicant_address_3), "Unknown", applicant_address_3)
  ) %>%
  ggplot(aes(x = applicant_address_3, y = inspection_result, fill = count_log)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "C", direction = -1, name = "Log(count)") +
  theme_bw(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  ) +
  labs(
    title = "Heatmap of Inspection Results Across Geographies",
    x = "Applicant Address (Geography)",
    y = "Inspection Result"
  )
```
## To answer question 3 we will use only Los Angeles data

```{r}
bldg_safty_insp_permit_2_LA = bldg_safty_insp_permit_2 %>% 
  filter(applicant_address_3 == "LOS ANGELES, CA") %>% 
  # remove duplicates
  distinct()
bldg_safty_insp_permit_2_LA %>%  dim()

bldg_safty_insp_permit_2_LA %>% head(15)

```
# Check if the naming of contractor city is unified
```{r}
bldg_safty_insp_permit_2_LA %>% 
  count(contractor_city, name = "count") %>% 
  arrange(desc(count)) %>% 
  print(n=100)

```
Yes mostly it is

# now convert the contractor data to 0 (not from Los Angeles) and 1 (from Los Angeles)
```{r}
bldg_safty_insp_permit_2_LA_2 = bldg_safty_insp_permit_2_LA %>%
  mutate(
    contractor_city_binary = case_when(
      str_detect(contractor_city, "LOS ANGELES") ~ 1,
      TRUE ~ 0
    )
  )

bldg_safty_insp_permit_2_LA_2 %>% head()
```


```{r}
# Check if the contractor city is converted to 0 and 1
table(bldg_safty_insp_permit_2_LA_2$contractor_city)
```

## Now check the types of inspection outcome
```{r}
inspection_result_count =
bldg_safty_insp_permit_2_LA_2 %>% 
  count(inspection_result, name = "count") %>% 
  arrange(desc(count)) %>% 
  filter (count > 1000) %>%
  print(n=100)

```
## Now we will select them and convert them to binary based on success 1 or no success 0
```{r}
bldg_safty_insp_permit_2_LA_3 = bldg_safty_insp_permit_2_LA_2 %>% 
  filter(inspection_result %in% inspection_result_count$inspection_result) %>%
           # remove duplicates
           distinct()
bldg_safty_insp_permit_2_LA_3 %>% dim()
bldg_safty_insp_permit_2_LA_3 %>% head(15)

## Now we will convert the inspection result to binary
bldg_safty_insp_permit_2_LA_3_insp_binary = bldg_safty_insp_permit_2_LA_3 %>%
  mutate(
    inspection_binary = case_when(
      inspection_result %in% c("Approved", "Permit Finaled", "CofO Issued", 
                               "OK for CofO", "OK to Issue CofO", "Completed", 
                               "SGSOV Approved") ~ 1,
      inspection_result %in% c("Insp Scheduled", "Partial Approval", 
                               "Not Ready for Inspection", "Corrections Issued",
                               "Partial Inspection", "Insp Cancelled", 
                               "Conditional Approval", "Cancelled", 
                               "No Access for Inspection", "SGSOV No Gas", 
                               "CofO in Progress", "CofO on Hold", 
                               "Off-Hour Fees Due", "Pending Review") ~ 0,
      TRUE ~ NA_real_  # Handles NA or unmatched values
    )
  )
bldg_safty_insp_permit_2_LA_3_insp_binary %>% dim()
bldg_safty_insp_permit_2_LA_3_insp_binary %>% head(15)
```

# apply chi square test to see if there is a relationship between inspection result and contractor city
```{r}
# Create a contingency table
contingency_table = table(bldg_safty_insp_permit_2_LA_3_insp_binary$contractor_city,
                           bldg_safty_insp_permit_2_LA_3_insp_binary$inspection_binary)
print(contingency_table)
```

```{r}
chi_test_result = chisq.test(contingency_table)
print(chi_test_result)
```
### It shows that the contractor from Los Angeles has a higher chance of getting a successful inspection result.
But it hightly depends on how the variables were converted to binary. It will need to be checked with the domain expert.
